---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo = TRUE}
unzip(zipfile = "activity.zip")
act <- read.csv("activity.csv", header = TRUE, sep=",",na.strings = "NA")
act <- transform(act, date=as.Date(as.character(date)))
```

## What is mean total number of steps taken per day?

Total number of steps taken per day is represented by spd
```{r, echo = TRUE}
spd <- with(act, tapply(steps, date, sum, na.rm = TRUE))
head(spd)
```

```{r, echo = TRUE}
hist(spd)
```

Mean and median of the total number of steps taken per day
```{r, echo = TRUE}
mean(spd)
median(spd)
```

## What is the average daily activity pattern?

```{r, echo = TRUE}
spi <- with(act, tapply(steps, interval, mean, na.rm=TRUE))
plot(names(spi),spi, type="l")
```


The maximum number of steps representing the 5-minute interval
```{r, echo = TRUE}
names(spi[which.max(spi)])
```
## Imputing missing values

The total number of missing values 
```{r, echo = TRUE}
length(which(is.na(act)))
```

Missing values are replaced with the mean for that 5-minute interval
```{r, echo = TRUE}
act[which(is.na(act$steps)),]$steps <- tapply(act$steps,act$interval,mean, na.rm=TRUE)
head(act)
```

```{r, echo = TRUE}
spd2 <- with(act, tapply(steps, date, sum, na.rm = TRUE))
hist(spd2)
mean(spd2)
median(spd2)
```
## Are there differences in activity patterns between weekdays and weekends?

Creating a new two levels factor variable
```{r, echo = TRUE}
wd <- as.factor(weekdays(act$date, abbreviate = TRUE))
levels(wd) <- c("weekday","weekday","weekend","weekend","weekday","weekday","weekday")
act <- cbind(act,wd)
d <- aggregate(act$steps, by = list(act$interval,act$wd), FUN="mean")
```

```{r, echo = TRUE}
library(lattice)
xyplot(x~Group.1|Group.2, data=d, layout=c(1,2), type="l")

```